---
title: "Cooking with GermaParl"
date: "2023-06-01"
output: ioslides_presentation
---

```{r libraries, echo = FALSE}
library(DiagrammeR)
library(xml2)
library(kableExtra)
library(polmineR)
library(dplyr)
library(purrr)
```


## Data formats and structures{.smaller}

### Data formats{.smaller}

* TEI-inspired XML: https://github.com/PolMine/GermaParlTEI
* Corpus Workbench (CWB): https://zenodo.org/record/7949074 (indexed and linguistically annotated)

---

### Data Preparation{.smaller}

```{r make_diagrammeR", echo = FALSE}
grViz("
  digraph {
  graph [layout = dot, rankdir = LR, nodesep = 0.5]
  

node [shape = cylinder style = filled fillcolor = Gray93]
RAW1[label = 'Raw Data 1']
RAW2[label = 'Raw Data 2']
RAW3[label = 'Raw Data 3']
XML[label = 'TEI/XML']
XML2[label = 'TEI/XML (enriched)']
VRT[label = 'VRT']
CWB[label = 'CWB']

node[shape = rect style=filled fillcolor = Gray]
PARSE[label = 'Preprocessing and Parsing\n(frappp)']
ENRICH[label = 'Enrich\n(frappp)']
ANNOTATE[label = 'Linguistic Annotation\n(bignlp/Stanford CoreNLP, TreeTagger)']
ENCODE[label = 'Encoding\n(Corpus Workbench)']


RAW1 -> PARSE
RAW2 -> PARSE
RAW3 -> PARSE

PARSE -> XML -> ENRICH-> XML2 -> ANNOTATE -> VRT -> ENCODE -> CWB
   }
")
```

---

## XML/TEI{.smaller}

* structural annotation (metadata on document and speaker level)
* quasi standardization, inspired by the standards of the Text Encoding Initiative

```{r read_xml, echo = FALSE}
xml2::read_xml(x = "~/Lab/github/GermaParlTEI/01/BT_01_003.xml") %>%
  as.character() %>%
  cat()
```

---

## Linguistic Annotation in the Verticalized Text format (VRT){.smaller}

* linguistically annotated "verticalized text" (tab separated and annotated with XML tags)
* see Corpus Encoding Manual (https://cwb.sourceforge.io/files/CWB_Encoding_Tutorial/)

```{r read_vrt, echo = FALSE}
xml2::read_xml("~/Lab/tmp/GermaParl2/vrt/BT_01_003.vrt") %>%
  as.character() %>%
  strsplit(split = "\n") %>% 
  pluck(1) %>% 
  head(26) %>% 
  cat(sep = "\n")
```

---

## Linguistic Annotation in GermaParl{.smaller}

### Positional attributes{.smaller}

Attributes at the token level are *positional attributes*:

* word: the "surface form" of the token
* xpos: Part-of-Speech-Annotation #1 (Stuttgart-Tübingen Tagset / STTS)
* upos: Part-of-Speech-Annotation #2 (Universal Dependencies Tagset)
* lemma: Lemmatized form of the token

```{r sample_ts, echo = FALSE}
sample_ts <- partition("GERMAPARL2", protocol_date = "1949-09-15") %>%
  polmineR::decode(
    s_attributes = character(), 
    p_attributes = c("word", "xpos", "upos", "lemma"),
    to = "data.table"
  ) %>%
  .[1:10]

knitr::kable(
  sample_ts, 
  format = ifelse(knitr::is_html_output(), "html", "latex"),
  booktabs = TRUE, 
  escape = TRUE, 
  caption = "A token stream representation of the GermaParl corpus") %>%
  footnote(general = "Beginning of session 1/3 (1949-09-15)")
```

---

### Structural attributes{.smaller}

Aside from annotation on token level, sentences, paragraps and named entities are linguistically annotated. These can comprise more than one token and thus are annotated as *structural attributes*. The relevant structural attributes are:

* `p` - regions defining paragraphs (no distinct values). 
* `p_type` distinguishes between speeches and other kinds of text such as stage comments or interjections
* `s` - regions defining sentences (no distinct values). 
* `ne_type` describes the type of a region which has been annotated as a named entity. These can be "ORGANIZATION", "LOCATION", "MISC" or "PERSON".

The `p` and `s` annotation layers can be used to subset the corpus based on sentences and paragraphs or to set boundaries for analyses like `kwic()`.

---

## Working with linguistic annotations {.smaller}

### CQP: The corpus query processor{.smaller}

`polmineR` exposes the powerful CQP query language to query CWB corpora. Using the CQP syntax, the linguistic annotation layers are accessible. The CWB development team provides a comprehensive handbook (https://cwb.sourceforge.io/files/CQP_Manual.pdf).

---

### Regular expressions

Simple regular queries work on the surface form of the token, i.e. the word...

```{r kwic_regular}
corpus("GERMAPARL2") |>
  kwic(query = "Transformation", s_attribute = "protocol_date")
```


---

### Accessing positional attributes

... the CQP syntax allows to specify the positional attribute like so:

```{r kwic_cqp_1}
corpus("GERMAPARL2") |>
  kwic(query = '[word = "Transformation"]', s_attribute = c("protocol_lp", "protocol_no"))
```

---

Note the single quotation marks around the entire query as well as the squared brackets.

This introduces some flexibility to the query. For example, the part of speech tag can be used as additional information. 

While the German word for "Fliege" is ambivalent (just like the English word) where it can describe an animal or a specific form of the verb "to fly", the CQP syntax allows to differentiate between the two. 

```{r kwic_comp_1}
corpus("GERMAPARL2") |>
  kwic(query = '[word = "Fliege" & xpos = "V.*"]', s_attribute = c("protocol_lp", "protocol_no"))
```

```{r kwic_comp_2}
corpus("GERMAPARL2") |>
  kwic(query = '[word = "Fliege" & xpos = "N.*"]', s_attribute = c("protocol_lp", "protocol_no"))
```

---

## CQP processor and advanced queries{.smaller}

CQP queries can be more complex and work for most `polmineR` methods.

```{r count_1}
corpus("GERMAPARL2") |>
  count(query = '[xpos = "AD.*"] [word = "Transformation"]',
        breakdown = TRUE)
```

The manual does provide some even more advanced queries. 

```{r count_2}
corpus("GERMAPARL2") |>
  subset(protocol_lp %in% c(5, 6)) |>
  count(query = '(?longest)[xpos = "V.*"] []{0,5} [word = "Transformation"]',
        breakdown = TRUE)
```

---

#### Linguistic Annotations in structural attributes{.smaller}

Linguistic annotations can be used in structural attributes as well.

```{r count_ner}
corpus("GERMAPARL2") |>
  polmineR::count(
    query = '/region[ne_type,a]::a.ne_type="ORGANIZATION"',
    cqp = TRUE,
    breakdown = TRUE
  )
```

---

It can be used to subset the corpus as well. In the following example, the type of the paragraph is used to include only the contents of actual speeches and not other stage comments.

```{r subset}
speech_only <- corpus("GERMAPARL2") |>
  subset(p_type == "speech") |>
  subset(speaker_name == "Angela Merkel")
```