---
title: "CWB manifesto corpus from scratch"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Problem statement

Getting pdf documents of manifestos, extracting text, cleaning the data and building a manifesto corpus would be a great Fingerübung. But the corpus of the Manifesto Project is one established corpus used by many researchers. As it is licensed data, a corpus built on it cannot be shared freely. But we can share the code to build the corpus, and this is what I want to do here.

```{r}
if (!"manifestoR" %in% unname(installed.packages()[,"Package"])){
  install.packages("manifestoR")
}
library(manifestoR)
library(tibble)
library(cwbtools)
library(data.table)
library(RcppCWB)
library(polmineR)
library(tokenizers)
```

To gain access to the Manifesto Database, you need an API key. 

```{r}
mp_setapikey("~/.credentials/manifesto_api_key.txt")
```


# downloading manifesto dataset

```{r}
mpds <- mp_maindataset()
de <- mp_corpus(countryname == "Germany")
```

This is a pretty condensed way to turn this into a tibble.

```{r}
tbl <- tibble(
  party = as.character(unname(sapply(lapply(de,  `[[`, "meta"), `[[`, "party"))),
  date = as.character(unname(sapply(lapply(de,  `[[`, "meta"), `[[`, "date"))),
  txt = unlist(sapply(lapply(lapply(de,  `[[`, "content"), `[[`, "text"), paste, collapse = "\n"))
)
```


```{r}
tok <- lapply(
  tbl$txt,
  function(txt){
    tokenize_words(
      x = unname(unlist(tokenize_sentences(x = txt))),
      lowercase = FALSE, strip_punct = FALSE
    )
  }
)
body <- lapply(tok, function(x) paste(sprintf("<s>\n%s\n</s>\n", sapply(x, paste, collapse = "\n")), collapse = "\n"))

xml <- paste(
  sprintf(
    "<text party='%s' date='%s'>\n%s\n</text>",
    tbl$party, tbl$date, body
  ),
  collapse = "\n"
)
which(xml == "<")
vrt_dir <- file.path(tempdir(), "vrt")
dir.create(vrt_dir)
cat(xml, file = file.path(vrt_dir, "manifestos.vrt"))
data_dir <- file.path(tempdir(), "data_dir")
dir.create(data_dir)
```

The new corpus still needs to be loaded.

```{r}
cwb_encode(
  corpus = "MANIFESTOS",
  registry = registry(),
  vrt_dir = vrt_dir,
  data_dir = data_dir,
  encoding = "utf8",
  p_attributes = "word",
  s_attributes = list(text = c("party", "date"), s = character()),
)

p_attr <- "word"
cwb_makeall(corpus = "MANIFESTOS", p_attribute = p_attr, registry = registry())
cwb_huffcode(corpus = "MANIFESTOS", p_attribute = p_attr, registry = registry())
cwb_compress_rdx(corpus = "MANIFESTOS", p_attribute = p_attr, registry = registry())
```

```{r}
cl_load_corpus(corpus = "MANIFESTOS", registry = registry())
cqp_load_corpus(corpus = "MANIFESTOS", registry = registry())
```

This is a rudimentary check (using low-level RcppCWB functions) whether to corpus can be used. How often does the token "pride" occur?

```{r}
count("MANIFESTOS", query = "Digitalisierung", )
```

```{r, render = knit_print}
kwic(
  "MANIFESTOS", query = "Krieg",
  s_attributes = c("text_party", "text_date"),
  left = c(s = 1L), right = c(s = 1L)
)
```


```{r}
mp_cite()
```