---
title: "CWB manifesto corpus from scratch"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Problem statement

Getting pdf documents of manifestos, extracting text, cleaning the data and building a manifesto corpus would be a great Fingerübung. But the corpus of the Manifesto Project is one established corpus used by many researchers. As it is licensed data, a corpus built on it cannot be shared freely. But we can share the code to build the corpus, and this is what I want to do here.

```{r}
if (!"manifestoR" %in% unname(installed.packages()[,"Package"])){
  install.packages("manifestoR")
}
library(manifestoR)
library(tibble)
library(cwbtools)
library(data.table)
library(RcppCWB)
library(polmineR)
```

To gain access to the Manifesto Database, you need an API key. 

```{r}
mp_setapikey("~/.credentials/manifesto_api_key.txt")
```


# downloading manifesto dataset

```{r}
mpds <- mp_maindataset()
de <- mp_corpus(countryname == "Germany")
```

This is a pretty condensed way to turn this into a tibble.

```{r}
tbl <- tibble(
  party = as.character(unname(sapply(lapply(de,  `[[`, "meta"), `[[`, "party"))),
  date = as.character(unname(sapply(lapply(de,  `[[`, "meta"), `[[`, "date"))),
  txt = unlist(sapply(lapply(lapply(de,  `[[`, "content"), `[[`, "text"), paste, collapse = "\n"))
)
```


```{r, message = FALSE, results = FALSE}
manifesto_data_dir <- file.path(tempdir(), "manifestos2")
if (!file.exists(manifesto_data_dir)) dir.create(manifesto_data_dir)
file.remove(list.files(manifesto_data_dir, full.names = TRUE))
```

... and instantiating a `CorpusData` object.

```{r}
Manifestos <- CorpusData$new()
```

The `tidytext` package offers an efficient workflow to create the token stream we can assign to the `CorpusData` object.

```{r}
tbl <- tidytext::unnest_tokens(tbl = tbl, output = "word", input = "txt", to_lower = FALSE)
Manifestos$tokenstream <- as.data.table(tbl)
```

We assign (zero-based!) corpus positions. 

```{r}
Manifestos$tokenstream[, cpos := 0L:(nrow(tbl) - 1L)]
```

Create table with structural attributes.

```{r}
cpos_max_min <- function(x) list(cpos_left = min(x[["cpos"]]), cpos_right = max(x[["cpos"]]))
Manifestos$metadata <- Manifestos$tokenstream[, cpos_max_min(.SD), by = c("party", "date")]
setcolorder(Manifestos$metadata, c("cpos_left", "cpos_right", "party", "date"))
Manifestos$metadata[, "party" := "SPD"]
```

Remove the metadata from the table in the `tokenstream` field and order the columns nicely. 

```{r}
Manifestos$tokenstream[, party := NULL][, date := NULL]
setcolorder(Manifestos$tokenstream, "cpos")
Manifestos$tokenstream
```

Ready to encode the corpus!

```{r, message = FALSE}
Manifestos$encode(
   corpus = "MANIFESTOS2", encoding = "utf8",
   p_attributes = "word", s_attributes = c("date", "party"),
   registry_dir = registry(), data_dir = manifesto_data_dir,
   method = "CWB", compress = TRUE
)
```

The new corpus still needs to be loaded.

```{r}
cl_load_corpus(corpus = "MANIFESTOS", registry = registry())
cqp_load_corpus(corpus = "MANIFESTOS", registry = registry())
```

This is a rudimentary check (using low-level RcppCWB functions) whether to corpus can be used. How often does the token "pride" occur?

```{r}
count("MANIFESTOS", query = "Digitalisierung")
```

```{r, render = knit_print}
context("MANIFESTOS", query = "Krieg") %>%
  kwic(s_attributes = "party")
```


```{r}
mp_cite()
```