---
title: "Cooking with GermaParl"
date: "2023-06-29"
output: 
  xaringan::moon_reader:
    css: ["./css/default.css", "./css/metropolis.css", "./css/robot-fonts.css", './css/polminify.css']
    nature:
      countIncrementalSlides: false
editor_options: 
  chunk_output_type: console
---

```{r libraries, echo = FALSE, message = FALSE, warning = FALSE}
library(polmineR)
```

# Purpose and Motivation

The creation of meaningful subsets or **subcorpora** of GermaParl enables researchers to
* focus on specific slices of the data 
* conduct diachronic analyses (changing patterns over time: dates, years, legislative periods)
* conduct synchronic analyses (difference between speakers and groups: speaker, parliamentary group, party, role)


But also focus on specific parts of the text
* speeches
* stage comments
* named entities

---

# Foundation: Structural Annotation

Subsetting is made possible by the *structural annotation* of the corpus.

Use `s_attributes()` to find out, which structural attributes are available.

```{r print_s_attributes}
s_attributes("GERMAPARL2")
```

To see which values these attributes can have, use the function with the attribute name.

```{r print_year}
s_attributes("GERMAPARL2", "protocol_year") |>
  head()
```

---

# Creating Subcorpora with subset()

In polmineR, a central function to create **subcorpora** is the `subset()` function.

To create a subcorpus object based on the initial corpus, select the structural attribute as well as its desired value. For example all debates in 1949:

```{r subset_year, eval = FALSE}
subset("GERMAPARL2", protocol_year == 1949)
```

**Note**: Passing the name of the corpus as a `character vector` is a shortcut. Passing a `corpus` object instead is more explicit.

```{r subset_year_pipe, eval = FALSE}
corpus("GERMAPARL2") |>
  subset(protocol_year == 1949)
```

`subset()` can be applied to a subcorpus as well. In addition, the creation of more complex slices of the corpus also works by chaining multiple calls to `subset()` as a "pipe":

```{r subset_pipe, eval = FALSE}
corpus("GERMAPARL2") |>
  subset(protocol_year == 1949) |>
  subset(speaker_parlgroup == "FDP")
```

---

# Creating Subcorpora with subset()

`subset()` allows for different additional ways to create a **subcorpus**, supporting non-standard evaluation.

Using `subset()` with `grepl()`:

```{r subset_grep}
germaparl_merkel <- corpus("GERMAPARL2") |>
  subset(grepl("Merkel", speaker_name))

s_attributes(germaparl_merkel, "speaker_name")
```

</br>

Using `subset()` with greater than, etc.

```{r subset_greater, eval = FALSE}
corpus("GERMAPARL2") |>
  subset(protocol_year >= 2000)
```

See the documentation of `polmineR` for further examples.

---

# Creating Subcorpus Bundles with split()

`split()` is used to split a corpus or subcorpus object into a bundle of subcorpora by a structural attribute.

For example, the entire corpus can be split into subcorpora representing legislative periods:

```{r split_1}
corpus("GERMAPARL2") |>
  split(s_attribute = "protocol_lp")
```


`split()` can also be used for subcorpora.

```{r split_subset, eval = FALSE}
corpus("GERMAPARL2") |>
  subset(protocol_lp == 19) |>
  split(s_attribute = "protocol_no")
```

---

It is also possible to split a subcorpus.

```{r example, eval = FALSE}
corpus("GERMAPARL2") |>
  subset(as.Date(protocol_date) >= as.Date("2020-01-01")) |>
  split(s_attribute = "protocol_date")
```

---

# as.speeches()

A natural unit of analysis for a corpus of plenary debates might be a single speech. A bundle of individual speeches can be created by `polmineR` with the special function `as.speeches()`.

```{r as_speeches, eval = FALSE}
corpus("GERMAPARL2") |>
  as.speeches(s_attribute_name = "speaker_name",
              s_attribute_date = "protocol_date")
```

Aside from `s_attribute_name` and `s_attribute_date`, `as.speeches()` also has an argument called `gap` which controls the number of tokens by which an utterance of the same speaker can be interrupted before it is split into two different speeches.

`as.speeches()` also works for subcorpora.

---

# Speeches and Interjections

Whether a token is part of a speech or part of an interjection can be an important distinction. In GermaParl v2, this information is part of the annotation of paragraphs.

```{r p_type, eval = FALSE}
corpus("GERMAPARL2") |>
  subset(p_type == "speech") |>
  subset(speaker_name == "Angela Merkel")
```

More technical details about the nested XML structure of the corpus are provided in the [Release Note](https://polmine.github.io/posts/2023/04/03/GermaParl-v2-beta3-Release-Note.html) of GermaParl v2 beta 3 where this and the next example are also discussed.

---

# Subsetting paragraphs and read()

With `read()` it is possible to access formatted full text of a subcorpus. To show the annotated interjections of GermaParl v2, the definition of paragraphs is necessary:

```{r read, eval = FALSE}
corpus("GERMAPARL2") |>
  subset(protocol_date == "2001-09-12") |>
  subset(p) |>
  read()
```

See the aforementioned  [Release Note](https://polmine.github.io/posts/2023/04/03/GermaParl-v2-beta3-Release-Note.html) for more technical details.